\textit{MSF: Modulated Sub-graph Finder}

### 2018-03-01 MSF_Paper_FA.tex
Update MSF_Paper_FA.tex

On these grounds we propose a novel approach to make use of the rich
pathway annotation resources available to gain additional functional
insights from basic DGEA. To this, we
start with the presupposition that expression of neighboring genes
within a functional pathway are not independent from each
other. Rather, they are often regulating each others expression or are
part of the same regulon~\cite{Michalak}. We
understand that the categorization of links between genes into labeled
pathways is often an arbitrary one, given the extensive cross talk
between different pathways. Although these categories have proven to be
useful in many situations, they force a certain perspective onto the
interpretation of novel data. Based on this two principles, we aim to
find sub-graphs within predefined networks which exhibit as a whole
significant differential expression changes. As input, information on
functional links between genes provided by e.g. KEGG or Reactome and
information on the differential expression status of single genes
resulting from a DGEA, are required. As a result the analysis returns
sub-graphs and their joint confidence scores, reflecting how the
perturbation is migrated through the network. Furthermore, the entry
points of perturbation in the networks and overlap with conventional
pathway categories are returned. All of this can be helpful to
understand the cause end effect of a stimulus and might inform about
potential points of intervention. The proposed algorithm was
implemented as a java program, which was named Modulated Sub-graph
Finder (\texttt{MSF}).

>>>>

On these grounds we propose a novel approach to make use of the rich
pathway annotation resources available to gain additional functional
insights from basic DGEA. To this, we
start with the presupposition that expression of neighboring genes
within a functional pathway are not independent from each
other. Rather, they are often regulating each others expression or are
part of the same regulon~\cite{Michalak}. We
understand that the categorization of links between genes into labeled
pathways is often an arbitrary one, given the extensive cross talk
between different pathways. Although these categories have proven to be
useful in many situations, they force a certain perspective onto the
interpretation of novel data. Based on these two principles, we aim to
find sub-graphs within predefined networks which exhibit as a whole
significant differential expression changes. As input, information on
functional links between genes provided by e.g. KEGG or Reactome and
information on the differential expression status of single genes
resulting from a DGEA, are required. As a result the analysis returns
sub-graphs and their joint confidence scores, reflecting how the
perturbation is migrated through the network. Furthermore, the entry
points of perturbation in the networks and overlap with conventional
pathway categories are returned. All of this can be helpful to
understand the cause end effect of a stimulus and might inform about
potential points of intervention. The proposed algorithm was
implemented as a java program, which was named Modulated Sub-graph
Finder (\texttt{MSF}).
